Introducing Statistical Design of Experiments to SPARQL Endpoint Evaluation

### 2013-07-24 exp1/iswc2013.tex
move wraptable
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@623 43ef1190-90fd-4dd7-a769-d0121f02b80e
factor/level explanation.  2 -> two
git-svn-id: https://svn.ifi.uio.no/repos/users/kjekje-phd@624 43ef1190-90fd-4dd7-a769-d0121f02b80e

First, note that the concept of parameters as considered above is
known as \emph{factors} in DoE. For each factor, one may make a
measurement of a \emph{response variable} at certain
\emph{levels}. For example, one may measure the response time of a
SPARQL endpoint by measuring it when it contains 1 or 2 million
triples. Levels are not constrained to be continous variables, they
can be discrete, or even two different instances of a class. The
ordering is not essential, but it may contribute to how easy it is to
understand the results, as we shall see. Then, the response time is
the response variable, the factor is the number of triples and there
are two levels, 1 or 2 million triples. We have chosen in this paper
to only deal with the relatively straightforward aspects of the DoE
formalism, and therefore constrain the experiment strictly to 2
levels.

@@

Similarly, \emph{interaction effects} for 2 factors $AB$ are defined
by comparing averages for equal versus non-equal levels of $A$ and
$B$.  Details of this theory, and how to compute the effects in
practice, using linear regression, are found in
\cite{wu2009experiments}.

@@

% latex table generated in R 2.15.1 by xtable 1.5-6 package
% Fri May 10 16:31:49 2013
% Modified by hand to remove 1's
\begin{wraptable}{l}{0.5\textwidth}
\caption{The magnitude of effects for some important main effects and interactions.}\label{tab:effects}
\medskip
\begin{tabular}{rr}
  \hline
Factors & Effect  \\ 
  \hline
Implement & $-$21.27 \\ 
  Implement:Optional & $-$11.49 \\ 
  Implement:Union & $-$6.21 \\ 
  Implement:TripleC:Lang1 & 4.12 \\ 
  TripleC:Lang & 4.20 \\ 
  Implement:TripleC & 4.32 \\ 
  Implement:BGPComp & 4.89 \\ 
  Lang:Union & 5.16 \\ 
  Implement:BGPComp:Lang & 5.16 \\ 
  BGPComp:Lang & 5.25 \\ 
  Implement:Lang:Union & 5.75 \\ 
  TripleC & 7.06 \\ 
  Implement:Lang & 7.73 \\ 
  Lang & 8.62 \\ 
  Optional & 11.30 \\ 
  Union & 17.18 \\ 
   \hline
\end{tabular}
\end{wraptable}

@@

We would like an overall test to see whether the new implementation is
better than the old, and the presence of inactive factors makes it
possible to average the performance of the new and the old
implementations into two distinct vectors. We do this by creating an
average over all the 32 level-combinations of the five active
environmental factors. This leaves us with 4 values for each of the
two levels of ``Implement'' and a two-sample t-test can be
performed. Effectively, we treat our experiment as a $2^6$ experiment
replicated 4 times.

>>>>

In the experiments DoE is concerned with, a \emph{response variable}
is measured under a various combinations of parameters.  These
parameters are known as \emph{factors}.  For each factor, a range of
possible values is fixed.  These values are known as \emph{levels} for
this factor.  Levels are not constrained to be continuous variables, they
can be discrete, or even two different instances of a class.
Experiments are run by choosing a combination of levels
for the factors.

For example, one may measure the execution time of a SPARQL endpoint
by measuring it when it contains 1 or 2 million triples.  Then, the
execution time is the response variable, the factor is the number of
triples and it has two levels, 1 or 2 million triples. We have chosen
in this paper to only deal with the relatively straightforward aspects
of the DoE formalism, and therefore constrain the experiment strictly
to two levels.

@@

Similarly, \emph{interaction effects} for two factors $AB$ are defined
by comparing averages for equal versus non-equal levels of $A$ and
$B$.  Details of this theory, and how to compute the effects in
practice, using linear regression, are found in
\cite{wu2009experiments}.

@@

@@

We would like an overall test to see whether the new implementation is
better than the old, and the presence of inactive factors makes it
possible to average the performance of the new and the old
implementations into two distinct vectors. We do this by creating an
average over all the 32 level-combinations of the five active
environmental factors. This leaves us with 4 values for each of the
two levels of ``Implement'' and a two-sample t-test can be
performed. Effectively, we treat our experiment as a $2^6$ experiment
replicated 4 times.
% latex table generated in R 2.15.1 by xtable 1.5-6 package
% Fri May 10 16:31:49 2013
% Modified by hand to remove 1's
\begin{wraptable}{l}{0.48\textwidth}
\caption{The magnitude of effects for some important main effects and interactions.}\label{tab:effects}
\bigskip\bigskip
\begin{tabular}{rr}
  \hline
Factors & Effect  \\ 
  \hline
Implement & $-$21.27 \\ 
  Implement:Optional & $-$11.49 \\ 
  Implement:Union & $-$6.21 \\ 
  Implement:TripleC:Lang1 & 4.12 \\ 
  TripleC:Lang & 4.20 \\ 
  Implement:TripleC & 4.32 \\ 
  Implement:BGPComp & 4.89 \\ 
  Lang:Union & 5.16 \\ 
  Implement:BGPComp:Lang & 5.16 \\ 
  BGPComp:Lang & 5.25 \\ 
  Implement:Lang:Union & 5.75 \\ 
  TripleC & 7.06 \\ 
  Implement:Lang & 7.73 \\ 
  Lang & 8.62 \\ 
  Optional & 11.30 \\ 
  Union & 17.18 \\ 
   \hline
\end{tabular}
\end{wraptable}
